---
title: "Homework 2"
author: "Jasmine Nakayama"
date: "February 14, 2018"
output:
  html_document: default
  word_document: default
---

Load packages:
```{r setup, message=FALSE}
library(car)
library(dplyr)
library(ggplot2)
```

#Question 1: What kind of R object is the Davis dataset?
```{r}
class(car::Davis)
```
#Question 2: How many observations are in the Davis dataset?
```{r}
car::Davis%>%
  summarize(nObs=n())
```
#Question 3: For reported weight, how many observations have a missing value?
```{r}
sum(is.na(car::Davis$repwt)) 
```
#Question 4: How many observations have no missing values?
```{r}
car::Davis%>%
filter(complete.cases(.))%>%
  summarize(n())
```
#Question 5: Create a subset containing only females. How many females are in this subset?
```{r}
#create subset with only females
females<-car::Davis%>% 
filter(sex=='F')
#count number in subset
summarize(females,n())
```

#Question 6: What is the average BMI for these individuals?
```{r}
#create new variable with height in meters
davishtmeter<-car::Davis%>%
mutate(htmeter=height/100)
#create new variable for BMI ("the ratio of the weight in kilograms divided by the square of the height in meters")"
davisbmi<-davishtmeter%>%
  mutate(bmi=weight/htmeter^2)
#obtain average BMI
davisbmi%>%
  summarize(mean(bmi))
```
#Question 7: How do these individuals fall into the BMI categories (what are the frequencies and relative %'s)?

Category    | BMI range (kg/m2)   
 ----------- | ------------------- 
 Underweight | <18.5              
 Normal      | 18.5 to <25              
 Overweight  | 25 to <30                
 Obese       | 30 or higher   
 
```{r}
#create new categories for BMI ranges
davisbmicat<-davisbmi%>%
  mutate(bmicat=if_else(bmi<18.5, "Underweight", 
                    if_else(bmi<25, "Normal", 
                      if_else(bmi<30, "Overweight", "Obese","Missing"),
                    "Missing"),
                "Missing"))
#get frequencies & relative percentages
davisbmicat%>%
 count(bmicat)

#not sure how to get relative percentages w/o janitor--LOOK UP LATER
```

#Question 8: Create a histogram of BMI. 
```{r}
davisbmicat%>%
ggplot()+geom_histogram(aes(bmi))
```

There is one outlier whose BMI is over 500.

#Question 9: Create side-by-side boxplots of the BMI distributions by gender
```{r}
#remove outlier
newdavisbmicat<-davisbmicat%>%
  filter(bmi<100)
#look at new histogram
newdavisbmicat%>%
ggplot()+geom_histogram(aes(bmi))
#create boxplot of BMI distribution divided by gender
newdavisbmicat%>%
ggplot(aes(x=sex, y=bmi, fill=bmi))+geom_boxplot()
```

#Question 10: Create a clustered bar chart of the BMI categories by gender
```{r}
newdavisbmicat%>%
ggplot(aes(x=bmicat, fill=sex))+geom_bar(position="dodge")
```