---
title: "HW52018"
author: "Jasmine Nakayama"
date: "3/10/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
Link to repository: https://github.com/jynakay/Assignments

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Download the dataset from the Canvas folder. Then let's load everything in:

```{r}

# Load all packages
library(phyloseq)
packageVersion("phyloseq")
library(ggplot2) 
packageVersion("ggplot2")
library(RColorBrewer)
packageVersion("RColorBrewer")

#Load HMP data
load("HMPv35.RData")
HMPv35
```
####Problem 1

Subset the HMPv35 object to obtain only the samples from Tongue_dorsum. Call this new object HMPv35sub2

```{r}
# Code to subset with new object HMPv35sub2
sub <- get_variable(HMPv35, "HMPbodysubsite") %in% c("Tongue_dorsum")
sample_data(HMPv35)$sub <- factor(sub)
HMPv35sub2 <- prune_samples(sample_data(HMPv35)$sub == TRUE, HMPv35)
summary(sample_data(HMPv35sub2))
HMPv35sub2
```

####Problem 2

Produce the geometric box plot of diversity measures for your object, HMPv35sub2
```{r}
#box plot for diversity here
alpha_meas = c("Observed", "Chao1", "ACE", "Shannon", "Simpson", "InvSimpson")
(p <- plot_richness(HMPv35sub2, "HMPbodysubsite", measures=alpha_meas))
p + geom_boxplot(data=p$data, aes(x=HMPbodysubsite, y=value, color=NULL), alpha=0.1)
```

####Problem 3

You see what taxa are most prevalent in your subset, HMPv35sub2

```{r}
#Calculate number of taxa in object
ntaxa(HMPv35sub2)

par(mar = c(10, 4, 4, 2) + 0.1) # make more room on bottom margin
N <- 20
barplot(sort(taxa_sums(HMPv35sub2), TRUE)[1:N]/nsamples(HMPv35sub2), las=2)

```

####Problem 4

Using your HMPv35sub2 object, throw the rare taxa out of that object, then reduce to only taxa in the phylum Bacteroidetes. Call this new object HMPv35sub2frbac 

```{r}

#throw out the rare taxa from the HMPv35sub2 object with new object HMPv35sub2frbac

#The next step filters out taxa with low occurrence throughout all samples
HMPv35subsub = filter_taxa(HMPv35sub2, function(x) sum(x > 3) > (0.2*length(x)), TRUE)

# The next step filters out all taxa that occur in less than .01% of samples
HMPv35subr  <- transform_sample_counts(HMPv35subsub, function(x) x / sum(x) )
HMPv35subfr <- filter_taxa(HMPv35subr, function(x) mean(x) > 1e-5, TRUE)
HMPv35subfr

#Finally subset to only bacteria in the phylum Bacteroidetes
HMPv35sub2frbac = subset_taxa(HMPv35subfr, Phylum=="Bacteroidetes")
HMPv35sub2frbac
```

####Problem 5
Using your HMPv35sub2frbac object, what is the distribution of the top 20 OTU's?

```{r}
#Calculate number of taxa in object
ntaxa(HMPv35sub2frbac)

par(mar = c(10, 4, 4, 2) + 0.1) # make more room on bottom margin
N <- 20
barplot(sort(taxa_sums(HMPv35sub2frbac), TRUE)[1:N]/nsamples(HMPv35sub2frbac), las=2)

```

####Problem 6

Plot sample abundance by body site for your object HMPv35sub2frbac

```{r}
#Pcode for the plot
title = "plot_bar; by site; Bacteroidetes only"
plot_bar(HMPv35sub2frbac, "HMPbodysubsite", "Abundance", title=title)

```


####Problem 7

You try it with your HMPv35sub2frbac object

```{r}

#code for your plot with family here
plot_bar(HMPv35sub2frbac, "HMPbodysubsite", "Abundance", "Phylum", title=title)

```


We can use the techniques of network science to illustrate how similar or distant samples are.

Using our filtered object with all phyla, we use the `plot_net` function to illustrate the "proximity" (or similarity) of samples, while denoting site and the sex of the participant.

```{r}
plot_net(HMPv35subfr, maxdist = 0.3, color = "HMPbodysubsite", shape="sex")
```

Let's redo switching which variable is colored and which variable is differentiated by shape.

```{r}
plot_net(HMPv35subfr, maxdist = 0.3, shape = "HMPbodysubsite", color="sex")
```
####Problem 8

Which method of display do you like best and why?

```{r}

#Place your answer here as another comment.
#I like the second graph more, because it is easier to distinguish sex by color rather than by shape. There is higher contrast between the differences.

```
