---
title: "Homework 01 Gapminder"
author: "Jasmine Nakayama"
date: "January 31, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---
Load packages:
```{r}
library(gapminder)
library(ggplot2)
```

### **Homework 01 Exercise - Task 1**
Modified to provide the standard deviation, median and sample size for life expectancy, set `digits=2`:

The standard deviation of life expectancy is `r round(sd(gapminder$lifeExp), digits=2)` years. 
The median life expectancy is `r round(median(gapminder$lifeExp), digits=2)` years. 
The sample size for life expectancy is `r round(length(gapminder$lifeExp), digits=2)` years.


### **Homework 01 Exercise - Task 2**

Modified r code chunk to provide the mean and standard deviation for life expectancy by continent:

```{r}
aggregate(lifeExp ~ continent, gapminder, mean)
```

```{r}
aggregate(lifeExp ~ continent, gapminder, sd)
```

### **Homework 01 Exercise - Task 3**

Produces box plot with jittered points of GDP per capita by continents:
```{r}
library(ggplot2)
ggplot(gapminder, aes(x = continent, y = gdpPercap)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```
Produces box plot with jittered points of population by continents:
```{r}
  library(ggplot2)
ggplot(gapminder, aes(x = continent, y = pop)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

### **Homework 01 Exercise - Task 4**

Produces bubble splot for life expectancy by GDP per cpaita for Asia and Americas in 1992. Size of bubbles represents population size:
```{r}
gap_with_colors <-
  data.frame(gapminder,
             cc = I(country_colors[match(gapminder$country,
                                         names(country_colors))]))
                                         
keepers <- with(gap_with_colors,
                continent %in% c("Americas","Asia") & year == 1992)
plot(lifeExp ~ gdpPercap, gap_with_colors,
     subset = keepers, log = "x", pch = 21,
     cex = sqrt(gap_with_colors$pop[keepers]/pi)/1500,
     bg = gap_with_colors$cc[keepers])
```

### **Homework 01 Exercise - Task 5**

Link to repository: [https://github.com/jynakay/N741Homework1](https://github.com/jynakay/N741Homework1).



